---
title: "Introduction to FragiliTidy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to FragiliTidy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`FragiliTidy` provides fast, tidyverse-friendly implementations of fragility
metrics for two-arm clinical trials:

- **Fragility Index (FI)** and **Reverse Fragility Index (RFI)** for
  dichotomous outcomes (Walsh et al., 2014), with custom 2x2 Fisher's exact
  and chi-squared routines and binary search.
- **Continuous Fragility Index (CFI)** for continuous outcomes
  (Caldwell et al., 2021), with summary-statistic simulation.
- **Reverse Continuous Fragility Index (rCFI)**: how many additional
  participants per arm would have been needed for a non-significant
  continuous outcome to reach significance.

```{r setup}
library(FragiliTidy)
library(dplyr)
library(tibble)
```

## Dichotomous outcomes

```{r}
trials <- tribble(
  ~study,    ~ie, ~ce, ~in_, ~cn,
  "Trial A",  10,  20,  100,  100,
  "Trial B",   5,  15,   80,   80,
  "Trial C",  30,  30,  200,  200
)

trials |>
  fragility_index(ie, ce, in_, cn) |>
  reverse_fragility_index(ie, ce, in_, cn)
```

## Continuous outcomes

### From raw data

```{r}
set.seed(1)
x <- rnorm(50, mean = 70, sd = 10)
y <- rnorm(50, mean = 50, sd = 10)
continuous_fragility_index_raw(x, y)
```

### From summary statistics

When raw per-patient outcomes are not available, the CFI is estimated by
simulating compatible datasets via rejection sampling and averaging the
iterative substitution result over `n_sim` simulations (Caldwell et al.,
2021):

```{r}
continuous_fragility_index_summary(
  intervention_mean = 70, control_mean = 50,
  intervention_sd   = 10, control_sd   = 10,
  intervention_n    = 100, control_n   = 100,
  seed = 1
)
```

`continuous_fragility_index_summary()` takes the same intervention/control
argument structure as [`fragility_index_vec()`], just with mean/SD/n in place
of event counts, and is likewise vectorised.

### Reverse CFI: distance from significance

For non-significant trials, the reverse CFI estimates how many additional
participants *per arm* would have been required to reach significance, given
the observed effect size and variance:

```{r}
reverse_continuous_fragility_index_summary(
  intervention_mean = 55, control_mean = 50,
  intervention_sd   = 10, control_sd   = 10,
  intervention_n    = 30, control_n    = 30,
  seed = 1
)
```

### Tidy interface

Both CFI variants integrate into `dplyr` pipelines:

```{r}
trials_continuous <- tribble(
  ~study,    ~intervention_mean, ~control_mean, ~intervention_sd, ~control_sd, ~intervention_n, ~control_n,
  "Trial X",  70,                 50,            10,               10,          50,              50,
  "Trial Y",  60,                 55,            15,               15,          40,              40
)

trials_continuous |>
  continuous_fragility_index(intervention_mean, control_mean, intervention_sd, control_sd, intervention_n, control_n) |>
  reverse_continuous_fragility_index(intervention_mean, control_mean, intervention_sd, control_sd, intervention_n, control_n)
```

`continuous_fragility_index_summary()` and `reverse_continuous_fragility_index_summary()`
are themselves vectorised, so they can also be called directly on data frame
columns inside `mutate()` or `case_when()`, just like `fragility_index_vec()`
and `reverse_fragility_index_vec()`:

```{r}
trials_continuous |>
  mutate(
    cfi = case_when(
      intervention_n >= 50 ~ continuous_fragility_index_summary(
        intervention_mean, control_mean, intervention_sd, control_sd,
        intervention_n, control_n, seed = 1
      ),
      TRUE ~ NA_real_
    )
  )
```

## References

- Walsh M, Srinathan SK, McAuley DF, et al. The statistical significance of
  randomized controlled trial results is frequently fragile: a case for a
  Fragility Index. *J Clin Epidemiol* 2014;67:622-628.
- Caldwell JE, Youssefzadeh K, Limpisvasti O. A method for calculating the
  fragility index of continuous outcomes. *J Clin Epidemiol* 2021;136:20-25.
- Hurley ET et al. The Continuous Fragility Index of Statistically
  Significant Findings in Randomized Controlled Trials in the Anterior
  Shoulder Instability Literature. *Am J Sports Med* 2023.
